\chapter*{Abstract}


COMET is a future high-precision experiment searching for charged lepton flavour
violation through the muon-to-electron conversion process. It aims to push the
intensity frontier of particle physics by coupling an intense muon beam with
cutting-edge detector technology. The first stage of the experiment, COMET
Phase\nobreakdash-I, is currently being assembled and will soon enter its data
acquisition period. It plans to achieve a single event sensitivity to $\mu$--$e$
conversion in aluminium of $3.1 \times 10^{-15}$.

This thesis presents a study of the sensitivity and backgrounds of \mbox{COMET
Phase\nobreakdash-I} using the latest Monte Carlo simulation data produced. The
background contribution from cosmic ray-induced atmospheric muons is estimated
using a backward Monte Carlo approach, which allows computational resources to
be focused on the most critical signal-mimicking events.

Analysis of a $\mu$--$e$ conversion simulation sample suggests that COMET
Phase\nobreakdash-I will reach a single event sensitivity of $3.6 \times
10^{-15}$ within 146 days of data acquisition. Our results suggest that, in that
period, on the order of $10^3$ atmospheric muons will enter the detector system
and produce an event similar enough to the conversion signal to pass all the
signal selection criteria. Most of these events will be rejected by the Cosmic
Ray Veto system, however, we expect at least 2.2 background events to sneak in
unnoticed. It is vital for the conversion search that these events be
discriminated from conversion electrons, for instance by using
Cherenkov threshold counters to distinguish between muons and electrons or,
alternatively, by developing a direction identification algorithm to reject some
fraction of the $\mu^+$-induced events.