\chapter*{Acknowledgements}

% Colleagues
I would like to thank my supervisor, Yoshi Uchida, for guiding me on this
journey. I am grateful for your teachings and the freedom I was given in
pursuing my objectives---sometimes from remote locations for extended periods of
time! Thank you also to the other members of the COMET group at Imperial, Per,
Phill, Kou and Roden, for the continuous feedback and constructive
conversations. 

Thanks to Ben and Ewen for the discussions that kickstarted my work on the COMET
experiment. I have lost count of how many times I went back to your theses to
find clarity.
To Valentin, thank you for the discussions concerning your original cosmic ray
study, from which mine is inspired in more ways than one.

And thank you to everyone from the COMET collaboration for undertaking such a
difficult project, striving to reach the best achievable outcome despite many
hurdles. It was a pleasure to discover some of your countries (and gastronomy),
and a shame that we missed a few opportunities to meet (and eat) because of the
global situation.


% Family & friends
To the members of my family, thank you for your patience and care. Florie, thank
you for proofreading this thesis, and well done managing to stay mostly awake in
the process. I am grateful, and pleasantly surprised, that you have accompanied
me throughout this adventure. Thanks also to all my close friends for the moral
support and the good times.

I dedicate this work to my mother and my grandfather.

