\chapter{Conclusion}

% Muon and LFV
Since the discovery of the muon in 1937, our understanding of its properties has
been steadily evolving. Today, although it fits nicely into the Standard Model
of particle physics, some questions remain standing, and recent tensions
observed between theory and experiment are suggesting that the SM does not give
a complete picture of the muon's underlying nature. Charged lepton flavour
violation is one of the ways in which Nature might deviate from the SM. If CLFV
were observed, it would be a clear indication of new physics. Neutrino-less
muon-to-electron conversion is one of the most sensitive channels to search for
CLFV as its signature is a mono-energetic electron with few sources of potential
backgrounds. According to theoretical constraints, this process ought to be
suppressed beyond experimental sensitivities. Therefore, any observation of the
$\mu$--$e$ conversion signal would yield clear evidence for CLFV and hence guide
the field toward a better understanding of physics beyond the Standard Model.

% COMET
The COMET Phase\nobreakdash-I experiment will soon start its data acquisition run toward the
search for muon-to-electron conversion in aluminium. It is expected to be a
hundred times more sensitive to $\mu$--$e$ conversion than the previous best
measurement by SINDRUM II. To achieve this sensitivity, COMET Phase\nobreakdash-I will
produce an intense pulsed muon beam directed toward an aluminium target to
create muonic aluminium atoms, from which conversion electrons may emerge. The
Cylindrical Detector, which surrounds the muon stopping target, is designed to
clearly identify conversion electrons while rejecting as many experimental
backgrounds as possible. 

% Software & sim
In order to cover the needs of the experiment in terms of simulation,
calibration, reconstruction, data formats and data analysis, the COMET
collaboration develops a comprehensive set of software utilities named ICEDUST.
The work presented throughout this thesis relies heavily on Monte Carlo (MC) data
produced by ICEDUST's {\sc Geant4} simulations of Phase\nobreakdash-I.
% Here is a good place to kind of introduce the problem of producing large-scale
% datasets and estimating atmospheric backgrounds
MC simulations allow us to study the experiment outcomes and to optimise its
design before assembly. However, they are computationally expensive and
typically only allow us to simulate a small fraction of the amount of data
expected to be collected. Additionally, traditional MC sampling is particularly
inefficient when the source of events is far from the detector system, as is the
case when considering cosmic ray-induced backgrounds. This thesis aims to
address these two limitations by partially circumventing the brute force MC
method.

% GAN
We investigate a novel approach to the mass-production of simulation data based
on Generative Adversarial Networks. We propose a neural network generator of
hits for the Cylindrical Drift Chamber, which can produce synthetic energy
deposits in the detector at a rate $10^6$ times higher than the ICEDUST
simulation. The machine learning model is trained on a sample of simulated hits
and learns from their features and relationships. The trained model allows us to
generate only a subset of the hits produced by MC simulation, but far more
efficiently than was previously achievable. Our ambition for this work and
future developments is to enable the production of entire mock datasets, which
can be used by the collaboration in anticipation of data acquisition runs.

% BMC & background study
Additionally, we present a study of the backgrounds caused by cosmic ray-induced
atmospheric muons in the COMET Phase\nobreakdash-I muon-to-electron conversion search. We
use a backward Monte Carlo simulation method to efficiently estimate the flux of
atmospheric muons near the detector system. We then perform a full analysis of
the event count contributions from $\mu$--$e$ conversion, muon decay-in-orbit,
and atmospheric muons. Our findings suggest that an efficient rejection of
atmospheric backgrounds by the Cosmic Ray Veto, combined with particle-type
identification by the Cylindrical Detector, are necessary for the Phase\nobreakdash-I search
to succeed. Under optimistic conditions, namely with a
\SI{99.99}{\percent}-efficient CRV and \SI{99}{\percent} muon identification
rate by the CyDet, we expect 0.08 atmospheric background events during the 146
days of data acquisition. According to our analysis, during this run time, COMET
Phase\nobreakdash-I will be able to achieve a single event sensitivity of $3.61 \times
10^{-15}$.


% Wrap up: what do my results mean for COMET!!!
Our most important contributions have been integrated into the main ICEDUST
software repository and documented for others to use and build upon. As the
COMET experiment goes into the Phase\nobreakdash-I measurement period, we hope that these
techniques will help to refine our understanding of the data and that they can
be applied in preparation for Phase\nobreakdash-II and beyond.